---
title: "Data Wrangling"
output: html_notebook
---

# SETUP
```{r include=FALSE}
library(tidyverse)

setwd('C:/Users/moth1/OneDrive/Documents/dsa-projects/2-shiny-app')
```

# LOAD FILES
```{r warning=FALSE, include=FALSE}
csv_filenames = list.files("./data", pattern = "*.csv")
csv_filepaths = paste('./data', csv_filenames, sep = '/')

data_length = length(csv_filepaths)
dflist = rep(list(1), data_length)

years = c(2014, 2016, 2017, 2018, 2019)
for (i in 1:data_length){
  df = read.csv(csv_filepaths[i], na.strings=c("n/a","N/A",""))
  year = data.frame(year = rep(years[i], dim(df)[1]))
  dflist[[i]] = bind_cols(year, df)
}
backup = dflist
```

# DATA WRANGLING

## Data Preparation

### Cleaning Variable Names (Questions)
```{r}
dflist = backup
clean.qs = function(df){
  qs = names(df)
  qs = lapply(qs, gsub, pattern='\\.strong\\.', replacement='.')
  qs = lapply(qs, gsub, pattern='\\.em\\.', replacement='.')
  qs = lapply(qs, gsub, pattern='\\._\\.', replacement='.')
  qs = lapply(qs, gsub, pattern='^X.', replacement='')
  qs = lapply(qs, gsub, pattern='\\.\\.\\.\\.', replacement='.')
  qs = lapply(qs, gsub, pattern='\\.\\.\\.', replacement='.')
  qs = lapply(qs, gsub, pattern='\\.\\.', replacement='.')
  names(df) = qs
  return(df)
}
dflist = lapply(dflist, clean.qs)
```

### Manual Pre-processing
```{r}
colnames(dflist[[1]])[3] <- "what.is.your.age."
colnames(dflist[[1]])[5] <- "what.country.do.you.work.in."
colnames(dflist[[1]])[10] <- "if.you.have.a.mental.health.issue.do.you.feel.that.it.interferes.with.your.work.when.NOT.being.treated.effectively."
#interpolate from other columns
dflist[[1]]["currently.have."] = ifelse((dflist[[1]][8] == "Yes")
                                        | (dflist[[1]][9] == "Yes"), "yes", "no")
qs = lapply(dflist, names)
```

### Functions to Verify Keywords/Questions
```{r}
verify = function(keyword) {
  cat(c("\n","\"", keyword, "\"", "\n"))
  cat("-----------------------\n")
  q.list <- grep(keyword, qs, ignore.case=TRUE)
  q.locs <- lapply(qs, grep, pattern=keyword, ignore.case=TRUE)  
  if (length(q.list)!=5){
    warning("Warning! Keyword not present in all years.")
    
  } else if (sum(unlist(lapply(q.locs, length)))>5){
    warning("Warning! Keyword present more than once in a year.")
    
  } else{
    warning("Keyword is a match!")
  }
  warning(" These are the matching questions:\n")
  print.qs(q.locs)
}
print.qs = function(q.locs){
  for (i in 1:5){
    for (j in 1:length(q.locs[[i]]))
    cat( c('List ',i,', ',
           'col ', q.locs[[i]][j], ': ',
           pluck(qs, i, q.locs[[i]][j]),
           '\n'), sep = "")
  }
}
```

### Manual Verification of Each Keyword/Questions
```{r}
# verify("year")
# verify("what.is.your.age.")
# verify("gender")
# verify("what.country.do.you.work.in.")
# verify("currently.have")
# verify("family.history.of.mental.illness.")
# verify("sought.treatment.")
# verify("interferes.with.your.work.when.not.")
# verify("how.many.employees.")
# verify("is.your.employer.primarily.a.tech.company.")
# verify("employer.provide.mental.health.benefits.")
# verify("do.you.know.the.options.for.mental.health.")
verify("disc.*a.mental.health.*.with.your.coworkers.")
```

## Data Loading

### Helpful Function to Find Variables Indixces
```{r}
keyword.locs = function(keyword, df) {
  qs = names(df)
  key.loc = grep(keyword, qs, ignore.case=TRUE)
  return(key.loc)
}
```

### Manually Define Keywords
```{r}
verified.qs = list("year",
                   "what.country.do.you.work.in.",
                   "what.is.your.age.",
                   "gender",
                   "currently.have.",
                   "is.your.employer.primarily.a.tech.company.",
                   "how.many.employees.",
                   "family.history.of.mental.illness.",
                   "sought.treatment.",
                   "interferes.with.your.work.when.not.",
                   "disc.*a.mental.health.*.with.your.coworkers.",
                   "employer.provide.mental.health.benefits.",
                   "do.you.know.the.options.for.mental.health.")
```

### Build Data Frame from Keywords
```{r}
df.tot = data.frame()

for (i in 1:data_length){   #for each data frame
  
  df = dflist[[i]]
  new.table = data.frame()
  
  for (j in 1:length(verified.qs)){   #for each question
    
    #unpack the question
    q = verified.qs[j]
    
    #find index in current table
    q.loc = keyword.locs(q, df)
    
    #extract question to new data frame
    new.col = select(df, q.loc)
    new.col[] = lapply(new.col, as.character)
    names(new.col) = tolower(q)
    
    #special case for j = 1
    if (j==1){
      new.table = new.col
    } else{
      new.table = bind_cols(new.table, new.col)
    }
  }
  #append to main data frame
  df.tot = bind_rows(df.tot, new.table)
}
df.tot = as.data.frame(sapply(df.tot, tolower))
backup2 = df.tot
```

### Pre-Data Cleaning EDA
```{r}
headers = names(df.tot)
for (i in 1:length(headers)){
  q = headers[[i]]
  barplot(table(df.tot[q]), main = q)
}
```

### Data Cleaning (Elements)
```{r}
df.tot = backup2

#rename columns
df.tot = df.tot %>% 
  rename(age = what.is.your.age.,
         country = what.country.do.you.work.in.,
         tech.company = is.your.employer.primarily.a.tech.company.,
         number.employees = how.many.employees.,
         mental.illness = currently.have.,
         family.history = family.history.of.mental.illness.,
         interferes.with.work = interferes.with.your.work.when.not.,
         comfortable.discussing = "disc.*a.mental.health.*.with.your.coworkers.",
         employer.provide.benefits = employer.provide.mental.health.benefits.,
         know.mental.health.options = do.you.know.the.options.for.mental.health.)

#handle missing data
df.tot = df.tot %>% 
  replace_na(list(age = median(df.tot$age, na.rm=TRUE),
                  gender = "non-binary",
                  tech.company = "yes",
                  number.employees = "self-employed"))

#handle redundant character responses
df.tot$country[df.tot$country=="united states"] = "united states of america"
df.tot = replace(df.tot, (df.tot==0) | (df.tot=="false") | (df.tot=="never"), "no")
df.tot = replace(df.tot, (df.tot==1) | (df.tot=="true") | (df.tot=="often") | (df.tot=="maybe")
                 | (df.tot=="rarely") | (df.tot=="sometimes") | (df.tot=="some of them"), "yes")
df.tot = map_df(df.tot, str_replace_all,
                pattern = "not eligible for coverage.*",
                replacement = "not eligible")
df.tot = map_df(df.tot, str_replace_all,
                pattern = "i don't know|don't know",
                replacement = "not sure")
df.tot = map_df(df.tot, str_replace_all,
                pattern = "i am not sure",
                replacement = "not sure")
df.tot = map_df(df.tot, str_replace_all,
                pattern = "not applicable",
                replacement = NA_character_)

#handle incorrect survey responses
df.tot$interferes.with.work[(df.tot$mental.illness=="yes") & (is.na(df.tot$interferes.with.work))] = "no"

gender = df.tot$gender
gender[str_detect(gender, "trans|fluid|queer|non|bi")] = "non-binary"
gender[str_detect(gender, "(?<!fe)male|^ma*.|^m$|dude")] = "male"
gender[str_detect(gender, "femal|woman|fem|^f$|she")] = "female"
gender[(gender!="male") & (gender!="female")] = "non-binary"
df.tot$gender = gender

#convert character variables to numeric/factors
df.tot$year = as.factor(df.tot$year)
df.tot$age = as.numeric(df.tot$age)
df.tot$age[(df.tot$age < 18) | (df.tot$age > 75)] = NA
df.tot$number.employees = factor(df.tot$number.employees,
                                 levels = c('self-employed','1-5','6-25','26-100','100-500','500-1000','more than 1000'),
                                 ordered = TRUE)
levels(df.tot$number.employees) = c('self-employed','1-25','1-25','26-100','101-500','>500','>500')
# df.tot$interferes.with.work = factor(df.tot$interferes.with.work,
#                                      levels = c('never','rarely','sometimes','often'), ordered = TRUE)

### reserved for filtering yes/no questions into only yes/no as.factor() ###


```

```{r}
# df.tot = backup
```

### Filters
```{r}
# df.tot %>% 
#   select(country) %>% 
#   group_by(country) %>% 
#   summarise(count = n()) %>% 
#   arrange(desc(count))

country.stat = df.tot %>% 
  select(country, sought.treatment) %>% 
  filter(country=="united states of america") %>% 
  filter(!is.na(sought.treatment)) %>% 
  summarise(percent.yes = sum(sought.treatment=="yes")/n())

global.stat =  df.tot %>% 
  select(sought.treatment) %>% 
  filter(!is.na(sought.treatment)) %>% 
  summarise(percent.yes = sum(sought.treatment=="yes")/n())

region = c("Global Survey", "United States")
percent.yes = c(global.stat, country.stat)
ggplot(data.frame(region,percent.yes)) + geom_col(aes(region, percent.yes, fill=region))
```


# DATA VISUALIZATION

### Post-Cleaning EDA Bar Charts
```{r}
headers = names(df.tot)
for (i in 1:length(headers)){
  q = headers[[i]]
  barplot(table(df.tot[q]), main = q)
}
```

### ggplot examples
```{r}
g = ggplot(df.tot)
g + geom_bar(aes(interferes.with.work, fill=tech.company), position="dodge")
g = df.tot %>% 
  filter(!is.na(interferes.with.work)) %>% 
  ggplot()
g + geom_bar(aes(tech.company, fill=interferes.with.work), position="fill")
df.tot$tech
```

# GRAPHS FOR PRESENTATION
```{r}
country.stat = df.tot %>% 
  select(country, interferes.with.work) %>% 
  filter(country=="india") %>% 
  filter(!is.na(interferes.with.work)) %>% 
  summarise(percent.yes = sum(interferes.with.work=="yes")/n())

global.stat = df.tot %>% 
  filter(!is.na(interferes.with.work)) %>% 
  summarise(percent.yes = sum(interferes.with.work=="yes")/n())

region = c("Global Survey", "India")
percent.yes = c(global.stat, country.stat)
ggplot(data.frame(region,percent.yes)) + geom_col(aes(region, percent.yes, fill=region)) +
  ggtitle("If you have a mental health condition,\n do you feel that it interferes with your work?") +
  ylab("Percent answered Yes") + xlab("")
```

```{r}
ggplot(filter(df.tot, !is.na(comfortable.discussing))) + geom_bar(aes(comfortable.discussing, fill=gender), position="fill") + ggtitle("Are you comfortable discussing mental health issues with a coworker?") + ylab("Percent of Respondents") + xlab("")
```

```{r}
year = c(2014:2017)
rate = c(13,13.3,13.4,14)
suicides = data.frame(year, rate)

x = df.tot %>% 
  filter(country=="united states of america") %>% 
  select(year, employer.provide.benefits) %>% 
  filter(!is.na(employer.provide.benefits)) %>% 
  group_by(year) %>% 
  summarise(percent.yes = sum((employer.provide.benefits=="yes")/n()))
x$year = as.numeric(as.character(x$year))
y = full_join(x,suicides)
ggplot(y) + geom_line(aes(x=year,y=percent.yes*20, group=1, color="Health Plan Coverage")) + geom_line(aes(x=year, y=rate, group=1, color="Natl. Suicide Rate")) + scale_y_continuous(name="Suicide Rate", sec.axis = sec_axis(~./20, name="Percent Covered")) + ggtitle("Mental Health Covered by Employer-Sponsored Health Insurance") + labs(color="Legend")

```


# OUTPUT FILES
```{r}
write_csv(df.tot, "server.csv")
```

